Unrecognized States: Theory and Cases

### 2017-07-01 theory.tex
first pass at cutting down section on sanctions

Let us begin with the simplest case, in which the sanctions affect only the secessionists' status quo payoffs, as when the imposition of sanctions has a negative impact on the economy of the unrecognized state. In the event that the secessionists cede or are defeated militarily, we presume the sanctions would be immediately lifted -- the home state would not want to sanction itself. In the event the secessionists achieve military victory, we presume they are able to force the home state to lift the sanctions.\footnote{Recall that our definition of military victory includes the ability to dictate the terms of settlement.} Sanctions are not expected to affect the payoffs to either ceding or military victory and neither should the cost of fighting itself be negatively impacted.

Thus, sanctions serve only two purposes: to narrow the difference between the payoffs from Status Quo and Fight, and to narrow the difference between the payoffs from Status Quo and Cede. If the patron wishes to maintain the status quo, its per-period investment must increase to compensate for the additional degradation of the status quo payoffs caused by the sanctions. All of this implies that the effect of sanctions on the unrecognized state's choice is not unambiguous. Proposition 2 lists necessary conditions for the existence of an equilibrium in which ``ceding'' is the equilibrium outcome once sanctions are introduced.

%Unfortunately, while the intended effect of sanctions is to reduce the unrecognized state's payoffs from the status quo, sanctions may also reduce the secessionists' payoffs from ceding.  When the home state collaborates with $c$ to enforce sanctions and impose economic suffering on the residents of the secessionist region, this can have the unintended consequence of increasing the hostility of the secessionists toward reunification.

@@

This logic is well illustrated by the case of Tamil Eelam, a territory in Northern Sri Lanka that existed as an unrecognized state from 1987-2009. Throughout the conflict, the Sri Lankan government used sanctions and blockades to reduce the secessionists' status quo payoffs and degrade their military capabilities. However, the effects of these sanctions were offset by two patrons: India, during the early stages of the conflict, and the Tamil diaspora throughout. The disruption of the status quo equilibrium began in 2006, when the United States, European Union, Canada, and India formally designated the leading secessionist organization, the Liberation Tigers of Tamil Eelam (LTTE), as terrorists.\footnote{Other states then followed suit, including Australia in 2008.} This designation greatly strengthened the sanctions against the secessionists and led to a sharp decline in both the quality of life in Tamil Eelam and the military capabilities of the LTTE. In January 2008, the Sri Lankan government abrogated the existing ceasefire agreement and in 2009 it launched a full-scale military offensive that ended with a decisive victory over the LTTE and reunification of Tamil Eelam and Sri Lanka. While there were many factors at play, the strengthening of sanctions increased the home state's probability of victory and thereby played a role in its decision to initiate a return to war.

%\vspace{3pt}
%\begin{center}
%\begin{tabular}{|  c| c | c | c |}
% \hline  $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
%	Cede& $$& $$ &$(-),(\downarrow)$ \\ \hline
%	Status Quo& $(\uparrow), (\downarrow)$& $(-),(\downarrow)$&$(-), (\downarrow)$ \\ \hline
%	Fight & $(\uparrow), (\downarrow)$ &$(\uparrow), (\downarrow)$&$(-),(\downarrow)$\\ \hline
% \multicolumn{4}{c} {\emph{Figure 5: Economic Sanctions by the Home State and Player $c$}}
%\end{tabular}
%\end{center}
%\vspace{5pt}

Despite this potential for perverse effects, sanctions are a common tool of outside actors ($c$), more common than aid and other positive inducements. A flow payment of carrots, even backed by the promises of a ``neutral" third party, may not be credible in the eyes of the secessionists, which could explain the frequent resort to sticks.

>>>>

Let us begin with the simplest case, in which the sanctions affect only the secessionists' status quo payoffs, as when the imposition of sanctions has a negative impact on the economy of the unrecognized state. We assume sanctions have only two effects: to narrow the difference between the payoffs from Status Quo and Fight, and to narrow the difference between the payoffs from Status Quo and Cede. If the patron wishes to maintain the status quo, its per-period investment must increase to compensate for the additional degradation of the status quo payoffs caused by the sanctions. In this case, the effect of sanctions on the unrecognized state's choice is ambiguous. Proposition 2 lists necessary conditions for the existence of an equilibrium in which ``ceding'' is the equilibrium outcome once sanctions are introduced.

@@
